Optimizing a PoS Tag Set for Dependency Parsing

### 2016-05-05 submission.tex
endringer i abstract

Our approach is somewhat semi-automatic in that we employ a "hybrid"
combination of manual annotation and automatic labeling in our experiments. Our
initial tags are gold standard, and using linguistic insight coupled with
computational considerations, we introduce new gold standard distinctions to
the tag set. The data sets with these tags are run through a pipeline with
automatically assigned tags in both training and testing, where the most
promising modifications are used "further". Hence, the tag set modifications
are not deterministic, nor do we try all possible combination; we only
experiment with linguistically motivated distinctions. We are in reality
altering the gold standard tags by the addition of more linguistic information.

@@

\paragraph{Parser}
In choosing a syntactic parser for our experiments, we considered previous work
on dependency parsing of Norwegian, specifically that of
\newcite{Sol:Skj:Ovr:14}, who found the Mate parser \cite{Boh:10} to be the
most successful parser for parsing of NDT. Furthermore, recent dependency
parser comparisons \cite{Cho:Tet:Ste:15} showed that Mate performed very well
on parsing of English, beating a range of contemporary state-of-the-art
parsers. Mate was consequently selected to evaluate the tag set modifications in
our experiments.

@@

\paragraph{Baseline}
It is common practice to compare the performance of PoS taggers to a
pre-computed baseline for an initial point of comparison.  For PoS tagging, a
commonly used baseline is the Most Frequent Tag (MFT) baseline, which we used
in our experiments. This involves labeling each word with the tag it was
assigned most frequently in the training. All unknown words, i.e., words not
seen in the training data, are assigned the tag most frequently assigned to
words seen only once in the training. Unknown and infrequent words have in
common that they rarely occur, and we might therefore expect them to have
similar properties.

\paragraph{Tags \& Features}
As we seek to quantify the effects of PoS tagging in a realistic setting, i.e.,
in application to raw text, we want to evaluate the parser on automatically
assigned PoS tags. For the training of the parser, however, we have two
options: using either gold standard or automatically assigned tags. In order to
settle on a configuration, we conducted experiments with gold standard and
automatically assigned tags to see how they differ with respect to performance.
The results of these experiments, shown in Table \ref{tagconfigeval}, reveal that
the combination of training and testing on automatic tags is clearly superior
to training on gold standard tags and testing on automatic tags. Consequently,
the parser was both trained and tested on automatically assigned tags in our
experiments.

@@

We removed any morphological features in order to simulate a realistic setting.
Moreover, it is crucial that we remove these features when working with
automatically assigned tags, as the features are still gold standard. For
instance, if a verb token is erroneously tagged as a noun, we could potentially
have a noun token with verbal features such as tense, which markedly obfuscates
the training and parsing. Another important aspect is that we want to isolate
the effect of PoS tags, necessitating the exclusion of morphological features.

>>>>

% fixme: tror nok dette bør flyttes og kortes ned. kan bakes inn med det som
% står under "Tag Set Mapping".
Our approach is somewhat semi-automatic in that we employ a "hybrid"
combination of manual annotation and automatic labeling in our experiments. Our
initial tags are gold standard, and using linguistic insight coupled with
computational considerations, we introduce new gold standard distinctions to
the tag set. The data sets with these tags are run through a pipeline with
automatically assigned tags in both training and testing, where the most
promising modifications are used "further". Hence, the tag set modifications
are not deterministic, nor do we try all possible combination; we only
experiment with linguistically motivated distinctions. We are in reality
altering the gold standard tags by the addition of more linguistic information.

@@

\paragraph{Parser}
In choosing a syntactic parser for our experiments, we considered previous work
on dependency parsing of Norwegian, specifically that of
\newcite{Sol:Skj:Ovr:14}, who found the Mate parser \cite{Boh:10} to have  
the best performance for parsing of NDT. Furthermore, recent dependency
parser comparisons \cite{Cho:Tet:Ste:15} showed that Mate performed very well
on parsing of English, outperforming a range of contemporary state-of-the-art
parsers. We will be using Mate for gauging the effect of the tag set modifications in
our experiments.

@@

\paragraph{Baseline}
As a point of reference for the PoS tagging we will be reporting the commonly used
most-frequent-tag baseline, henceforth MFT.
This involves labeling each word with the tag it was
assigned most frequently in the training data. All unknown words, i.e., words not
seen in the training data, are assigned the tag most frequently assigned to
words seen only once.  The rationale is that we might expect unknown
and infrequent words to have similar properties given that they both occur rarely.

\paragraph{Tags \& Features} % <-- fixme: rename? 
As we seek to quantify the effects of PoS tagging in a realistic setting, i.e.,
in application to raw text, we choose to evaluate the parser on automatically
assigned PoS tags. For training, however, we have two
options: using either gold standard or automatically assigned tags. In order to
settle on a configuration, we conducted experiments with training on both gold
standard tags and
automatically assigned tags to see how the performance differs.
The results of these experiments, shown in Table \ref{tagconfigeval}, reveal that
the combination of training and testing on automatic tags is clearly superior
to training on gold standard tags and testing on automatic tags. Consequently,
the parser is both trained and tested on automatically assigned tags in the
remainder of our experiments.

@@

Though some parsers make use of morphological features, % fixme: status for mate?
we removed all morphological features beyond the PoS tags in order to simulate a realistic setting.
Moreover, it is crucial that we remove these features when working with
automatically assigned tags, as such features would be provided by the gold
standard.

%For instance, if a verb token is erroneously tagged as a noun, we could potentially
%have a noun token with verbal features such as tense, which markedly obfuscates
%the training and parsing. Another important aspect is that we want to isolate
%the effect of PoS tags, necessitating the exclusion of morphological features.
